---
title: "Factbook Project"
author: "Vijay Kadiyala"
output: pdf_document
---

```{r,warning=FALSE}
library(XML)
```

```{r}
xmlfactbook <- xmlParse("factbook.xml")
```

#Homework Questions

##Homework 1

####Question 1

```{r}
infantmortality <- getNodeSet(xmlfactbook,'//field[@id="f2091"]/rank')
```


```{r}
countryvec <- NULL
mortality <- NULL
for (i in infantmortality){
  country <- xmlGetAttr(i,"country")
  countryvec <- append(countryvec,country)
}
for (i in infantmortality){
  mort <- xmlGetAttr(i,"number")
  mort <- as.numeric(mort)
  mortality <- append(mortality,mort)
}
infantmortalityDF <- data.frame(countryvec,mortality,stringsAsFactors = FALSE)
names(infantmortalityDF) <- c("cia","mortality rate")
#infantmortalityDF
```


####Question 2

```{r}
population <- getNodeSet(xmlfactbook,'//field[@id="f2119"]/rank')
#population
```

```{r}
popvec <- NULL
popcountryvec <- NULL
for (i in population){
  cnt <- xmlGetAttr(i,"country")
  popcountryvec <- append(popcountryvec,cnt)
}
for (i in population){
  pop <- xmlGetAttr(i,"number")
  pop <- as.numeric(pop)
  popvec <- append(popvec,pop)
}
populationDF <- data.frame(popcountryvec,popvec,stringsAsFactors = FALSE)
names(populationDF) <- c("cia", "population")
#populationDF
```

##Homework 2

####Question 1

```{r}
countrynames <- getNodeSet(xmlfactbook,'//appendix[@name="cross-reference list of country data codes"]//row')
cianames <- sapply(countrynames,function(x) xmlChildren(x)[1]$cell)
countryvec <- sapply(cianames,function(x) xmlGetAttr(x,"content"))
ciacode <- sapply(cianames,function(x) xmlGetAttr(x, "country"))
isocode <- sapply(countrynames,function(x) xmlGetAttr(xmlChildren(x)[3]$cell,"content"))
countryDF <- data.frame(countryvec,ciacode,isocode)
names(countryDF) <- c("country", "cia", "iso3166")
#countryDF
```

##Homework 3

```{r}
geocsv <- read.csv('countries_codes_and_coordinates.csv',stringsAsFactors = FALSE)
```


```{r}
geocsv <- geocsv[order(geocsv$Numeric.code),]
geocsv <- geocsv[!duplicated(geocsv$Numeric.code),]
geocsv$Alpha.2.code <- gsub(" ", "",geocsv$Alpha.2.code)
names(geocsv)[2] <- "iso3166"
#geocsv
```

##Homework 4

```{r}
library(dplyr)
homework4DF <- suppressWarnings(geocsv %>% inner_join(countryDF) %>% inner_join(populationDF) %>% inner_join(infantmortalityDF))
homework4DF <- homework4DF[c(7,2,8,9,10,5,6)]
#homework4DF

```

##Homework 7


```{r}
xmatrix <- as.matrix(homework4DF[5:7])
xmatrix <- scale(xmatrix)
xmatrix <- unname(xmatrix)
```

```{r}
centroid1 <- xmatrix[sample(1:nrow(xmatrix), 3, replace=FALSE),][1,]
centroid2 <- xmatrix[sample(1:nrow(xmatrix), 3, replace=FALSE),][2,]
centroid3 <- xmatrix[sample(1:nrow(xmatrix), 3, replace=FALSE),][3,]
```


```{r}
outputvector <- NULL
for(i in 1:nrow(xmatrix)){
  distance1 <- sqrt(sum((xmatrix[i,]-centroid1)^2))
  distance2 <- sqrt(sum((xmatrix[i,]-centroid2)^2))
  distance3 <- sqrt(sum((xmatrix[i,]-centroid3)^2))
  if (min(distance1,distance2,distance3) == distance1){
    outputvector <- append(outputvector,1)
  }
  if (min(distance1,distance2,distance3) == distance2){
    outputvector <- append(outputvector,2)
  }
  if (min(distance1,distance2,distance3) == distance3){
    outputvector <- append(outputvector,3)
  }
 
}
length(outputvector)
nrow(xmatrix)
```

```{r}
#Centroider subfunction: calculates centroids from X and output vector

centroider <- function(matrix,vector,k){
  centroids <- vector("list", length = k)
  for (i in 1:length(centroids)){
    group <- matrix[which(vector == i),]
    centroidvec <- NULL
    for (col in 1:ncol(group)){
      centroidvec <- append(centroidvec,mean(group[,col]))
    }
    centroids[[i]] <- centroidvec
    }
  return(centroids)
}
examplecentroids <- centroider(xmatrix,outputvector,k=3)
```


```{r}
#vectorizer subfunction: creates grouping vector given centroids
vectorizer <- function(matrix,centroidlist){
  finalvec <- NULL
  for (i in 1:nrow(matrix)){
    distancevec <- NULL
    for (centroid in centroidlist){
      distancevec <- append(distancevec,sqrt(sum((matrix[i,]-centroid)^2)))
    }
    finalvec <- append(finalvec,which(distancevec == min(distancevec)))
  }
  return(finalvec)
}
vectorizer(xmatrix,examplecentroids)
```

```{r}
#firstcentroid subfunction: randomly selects three points as initial centroids
firstcentroid <- function(matrix,k){
  centroids <- vector("list", length = k)
  sample <- matrix[sample(1:nrow(xmatrix), k, replace=FALSE),]
  for (i in 1:length(centroids)){
    centroids[[i]] <- sample[i,]
  }
  return(centroids)
}
firstcentroid(xmatrix,4)
```

```{r}

initialcentroid <- firstcentroid(xmatrix,3)
oldvec <- vectorizer(xmatrix,initialcentroid)
suppressWarnings(repeat{
  centroidfirst <- centroider(xmatrix,oldvec,k=3)
  newvec <- vectorizer(xmatrix,centroidfirst)
  if(newvec == oldvec){
    print(newvec)
    break()
  }
  oldvec <- newvec
  newvec <- NULL
})

```

```{r}
#final kmeans function: does kmeans clustering
kmeans <- function(matrix,k){
  matrix <- scale(matrix)
  matrix <- unname(matrix)
  initialcentroid <- firstcentroid(matrix=matrix,k=k)
  oldvec <- vectorizer(matrix=matrix,initialcentroid)
  suppressWarnings(
    repeat{
      centroid <- centroider(matrix,oldvec,k=k)
      newvec <- vectorizer(matrix,centroid)
      if(newvec == oldvec){
        return(newvec)
        break()
      }
      oldvec <- newvec
      newvec <- NULL
    }
  )
}
testmatrix <- as.matrix(homework4DF[5:7]) 
kmeans(testmatrix,k=6)
```

#Deliverables

####Deliverable 1

```{r}
hist(infantmortalityDF$`mortality rate`)
```

The distributon of infant mortality rates is right skewed. The majority of mortality rates seem to fall below 30, and very many countries have rates less than 10. There are a few countries with rates greater than 80. 

####Deliverable 2

```{r}
countryDF[countryDF$country == "Tajikistan",]
```

####Deliverable 3

a. 

```{r}
oldcsv <- read.csv('countries_codes_and_coordinates.csv',stringsAsFactors = FALSE)
#oldcsv[order(oldcsv$Numeric.code),]
#oldcsv[oldcsv$Numeric.code %in% which(duplicated(oldcsv$Numeric.code)),]
#which(duplicated(oldcsv$Numeric.code))
#oldcsv[duplicated(oldcsv$Numeric.code),]
oldcsv[duplicated(oldcsv$Numeric.code),]
```

b. 

```{r}
nrow(geocsv)
```

####Deliverable 4




####Deliverable 5

```{r}
library(dplyr)
homework4DF %>% filter(population > 50000000) %>% pull(`mortality rate`) %>% mean()
homework4DF %>% filter(population < 10000000) %>% pull(`mortality rate`) %>% mean()
```

####Deliverable 6

```{r}
breaks <- quantile(homework4DF$`mortality rate`,probs = seq(0,1,0.125))
breaks[1] <- 0
breakcuts <- cut(homework4DF$`mortality rate`,breaks)
table(breakcuts)
breaks
```

####Deliverable 7

```{r}
library(RColorBrewer)
#display.brewer.all()
mortalitycolors <- brewer.pal(8,'YlOrRd')

```

```{r}
install.packages('maps', repos = "http://cran.us.r-project.org")
```

```{r}
breakcolorsDF <- data.frame(table(breakcuts),mortalitycolors)
```

```{r}
"library(maps)
afghancolor <- breakcolorsDF[breakcolorsDF$breakcuts == breakcuts[41],]$mortalitycolors
afghancolor <- as.character(afghancolor)
map(world, fill = TRUE, col = grey)
symbols(homework4DF$Longitude..average.[41],homework4DF$Latitude..average.[41],circles=0.0005*sqrt(homework4DF$population[41]),add=TRUE,inches = FALSE,fg=afghancolor,bg=afghancolor)
"
```

```{r}
map("world", fill = TRUE, col = "grey")
for (i in 1:length(breakcuts)){
  color <- breakcolorsDF[breakcolorsDF$breakcuts == breakcuts[i],]$mortalitycolors
  color <- as.character(color)
  symbols(homework4DF$Longitude..average.[i],homework4DF$Latitude..average.[i],circles=4,add=TRUE,inches = FALSE,fg=color,bg=color)
}
```

####Deliverable 8

```{r}
map("world", fill = TRUE, col = "grey")
for (i in 1:length(breakcuts)){
  color <- breakcolorsDF[breakcolorsDF$breakcuts == breakcuts[i],]$mortalitycolors
  color <- as.character(color)
  radius <- 0.0005*sqrt(homework4DF$population[i])
  if (radius < 2){
    radius <- 2
  }
  symbols(homework4DF$Longitude..average.[i],homework4DF$Latitude..average.[i],circles=radius,add=TRUE,inches = FALSE,fg=color,bg=color)
}
```


```{r}
i<-1

repeat{

    print(i)

    if( i == 15) break

    i<-i+1

           }
```
