---
title: "Factbook Project"
author: "Vijay Kadiyala"
output: pdf_document
---

```{r,warning=FALSE}
library(XML)
```

```{r}
xmlfactbook <- xmlParse("factbook.xml")
```

#Homework Questions

##Homework 1

####Question 1

```{r}
infantmortality <- getNodeSet(xmlfactbook,'//field[@id="f2091"]/rank')
```


```{r}
countryvec <- NULL
mortality <- NULL
for (i in infantmortality){
  country <- xmlGetAttr(i,"country")
  countryvec <- append(countryvec,country)
}
for (i in infantmortality){
  mort <- xmlGetAttr(i,"number")
  mort <- as.numeric(mort)
  mortality <- append(mortality,mort)
}
infantmortalityDF <- data.frame(countryvec,mortality,stringsAsFactors = FALSE)
names(infantmortalityDF) <- c("cia","mortality rate")
#infantmortalityDF
```


####Question 2

```{r}
population <- getNodeSet(xmlfactbook,'//field[@id="f2119"]/rank')
#population
```

```{r}
popvec <- NULL
popcountryvec <- NULL
for (i in population){
  cnt <- xmlGetAttr(i,"country")
  popcountryvec <- append(popcountryvec,cnt)
}
for (i in population){
  pop <- xmlGetAttr(i,"number")
  pop <- as.numeric(pop)
  popvec <- append(popvec,pop)
}
populationDF <- data.frame(popcountryvec,popvec,stringsAsFactors = FALSE)
names(populationDF) <- c("cia", "population")
#populationDF
```

##Homework 2

####Question 1

```{r}
countrynames <- getNodeSet(xmlfactbook,'//appendix[@name="cross-reference list of country data codes"]//row')
cianames <- sapply(countrynames,function(x) xmlChildren(x)[1]$cell)
countryvec <- sapply(cianames,function(x) xmlGetAttr(x,"content"))
ciacode <- sapply(cianames,function(x) xmlGetAttr(x, "country"))
isocode <- sapply(countrynames,function(x) xmlGetAttr(xmlChildren(x)[3]$cell,"content"))
countryDF <- data.frame(countryvec,ciacode,isocode)
names(countryDF) <- c("country", "cia", "iso3166")
#countryDF
```

##Homework 3

```{r}
geocsv <- read.csv('countries_codes_and_coordinates.csv',stringsAsFactors = FALSE)
```


```{r}
geocsv <- geocsv[order(geocsv$Numeric.code),]
geocsv <- geocsv[!duplicated(geocsv$Numeric.code),]
geocsv$Alpha.2.code <- gsub(" ", "",geocsv$Alpha.2.code)
names(geocsv)[2] <- "iso3166"
#geocsv
```

##Homework 4

```{r}
library(dplyr)
homework4DF <- suppressWarnings(geocsv %>% inner_join(countryDF) %>% inner_join(populationDF) %>% inner_join(infantmortalityDF))
homework4DF <- homework4DF[c(7,2,8,9,10,5,6)]
#homework4DF

```


#Deliverables

####Deliverable 1

```{r}
hist(infantmortalityDF$`mortality rate`)
```

The distributon of infant mortality rates is right skewed. The majority of mortality rates seem to fall below 30, and very many countries have rates less than 10. There are a few countries with rates greater than 80. 

####Deliverable 2

```{r}
countryDF[countryDF$country == "Tajikistan",]
```

####Deliverable 3

a. 

```{r}
oldcsv <- read.csv('countries_codes_and_coordinates.csv',stringsAsFactors = FALSE)
#oldcsv[order(oldcsv$Numeric.code),]
#oldcsv[oldcsv$Numeric.code %in% which(duplicated(oldcsv$Numeric.code)),]
#which(duplicated(oldcsv$Numeric.code))
#oldcsv[duplicated(oldcsv$Numeric.code),]
oldcsv[duplicated(oldcsv$Numeric.code),]
```

b. 

```{r}
nrow(geocsv)
```

####Deliverable 4




####Deliverable 5

```{r}
library(dplyr)
homework4DF %>% filter(population > 50000000) %>% pull(`mortality rate`) %>% mean()
homework4DF %>% filter(population < 10000000) %>% pull(`mortality rate`) %>% mean()
```

####Deliverable 6

```{r}
breaks <- quantile(homework4DF$`mortality rate`,probs = seq(0,1,0.125))
breaks[1] <- 0
breakcuts <- cut(homework4DF$`mortality rate`,breaks)
table(breakcuts)
breaks
```

####Deliverable 7

```{r}
library(RColorBrewer)
#display.brewer.all()
mortalitycolors <- brewer.pal(8,'YlOrRd')

```

```{r}
install.packages('maps', repos = "http://cran.us.r-project.org")
```

```{r}
breakcolorsDF <- data.frame(table(breakcuts),mortalitycolors)
```

```{r}
"library(maps)
afghancolor <- breakcolorsDF[breakcolorsDF$breakcuts == breakcuts[41],]$mortalitycolors
afghancolor <- as.character(afghancolor)
map(world, fill = TRUE, col = grey)
symbols(homework4DF$Longitude..average.[41],homework4DF$Latitude..average.[41],circles=0.0005*sqrt(homework4DF$population[41]),add=TRUE,inches = FALSE,fg=afghancolor,bg=afghancolor)
"
```

```{r}
map("world", fill = TRUE, col = "grey")
for (i in 1:length(breakcuts)){
  color <- breakcolorsDF[breakcolorsDF$breakcuts == breakcuts[i],]$mortalitycolors
  color <- as.character(color)
  symbols(homework4DF$Longitude..average.[i],homework4DF$Latitude..average.[i],circles=4,add=TRUE,inches = FALSE,fg=color,bg=color)
}
```

####Deliverable 8

```{r}
map("world", fill = TRUE, col = "grey")
for (i in 1:length(breakcuts)){
  color <- breakcolorsDF[breakcolorsDF$breakcuts == breakcuts[i],]$mortalitycolors
  color <- as.character(color)
  radius <- 0.0005*sqrt(homework4DF$population[i])
  if (radius < 2){
    radius <- 2
  }
  symbols(homework4DF$Longitude..average.[i],homework4DF$Latitude..average.[i],circles=radius,add=TRUE,inches = FALSE,fg=color,bg=color)
}
```