---
title: "Project 4"
author: "Vijay Kadiyala"
output: pdf_document
---

```{r}
directory <- list.files("messages")
directory
```


#Homework

##Homework 1

```{r}
#This is a sample email for example purposes
easyhamfiles <- list.files("messages/easy_ham", full.names = TRUE)
hardhamfiles <- list.files("messages/hard_ham",full.names=TRUE)
easy2hamfiles <- list.files("messages/easy_ham_2",full.names=TRUE)
spam2files <- list.files("messages/spam_2",full.names=TRUE)
ham3 <- readLines(easyhamfiles[3])
ham69 <- readLines(easyhamfiles[69])
ham424 <- readLines(easyhamfiles[424])
ham77 <- readLines(easyhamfiles[77])
ham4 <- readLines(easyhamfiles[4])
hardham40 <- readLines(hardhamfiles[40])
hardham270 <- readLines(hardhamfiles[270])
```



```{r}
#Spamfiles 
spamfiles <- list.files("messages/spam", full.names = TRUE)
```


```{r}
splitMessage <- function(email){
  blanks <- email == ""
  bodysplit <- min(which(blanks == TRUE))
  header <- email[1:(bodysplit-1)]
  body <- email[(bodysplit+1):length(email)]
  outputlist <- list(header,body)
  names(outputlist) <- c("header","body")
  return(outputlist)
}
ham3split <- splitMessage(ham3)
ham69split <- splitMessage(ham69)
ham424split <- splitMessage(ham424)
ham77split <- splitMessage(ham77)
ham4split <- splitMessage(ham4)
hardham270split <- splitMessage(hardham40)
```

##Homework 2

```{r}
splitMessage(readLines(spamfiles[4]))$body
```

```{r}
hasAttachment <- function(header){
  contentlocations <- grepl("Content-Type",header)
  if ((TRUE %in% grepl("Content-Type",ham3split$header)) == FALSE){
    return(FALSE)
  }
  content_type_line <- header[min(which(contentlocations == TRUE))]
  content_type_line <- tolower(content_type_line)
  return(grepl("multipart",content_type_line))
}
#hasAttachment(ham77split$header)
hasAttachment(splitMessage(readLines(spamfiles[900]))$header)

```


```{r}
#content_line <- ham424split$header[grepl("boundary=",ham424split$header,ignore.case = TRUE)]
#strsplit(content_line,'=')
ham424boundary <- grep("boundary=",ham424split$header,value=TRUE,ignore.case = TRUE)
```



```{r}
getBoundary <- function(header){
  boundary <- grep("boundary=",header,value=TRUE,ignore.case=TRUE)
  boundaryname <- strsplit(boundary,"boundary=")[[1]][2]
  boundaryname <- noquote(boundaryname)
  boundaryname <- gsub(';','',boundaryname)
  boundaryname <- gsub(' ','',boundaryname)
  boundaryname <- gsub('\"','',boundaryname)
  return(boundaryname)
}
#splitMessage(readLines(spamfiles[900]))$header
#getBoundary(hardham40split$header)
```



```{r}
extractBodyText <- function(email){
  #email <- readLines(email)
  emailheader <- splitMessage(email=email)$header
  emailbody <- splitMessage(email=email)$body
  if (hasAttachment(emailheader) == FALSE){
    return(splitMessage(email=email)$body)
  }
  if (hasAttachment(emailheader)){
    boundary <- getBoundary(emailheader)
    boundarymatches <- grep(boundary,emailbody,fixed=TRUE)
    if (length(boundarymatches)==1){
      return(emailbody[(boundarymatches[1]+1):length(emailbody)])
    }
    if (length(boundarymatches) > 1){
      return(emailbody[(boundarymatches[1]+1):(boundarymatches[2]-1)])
    }
  }
  return(0)
}
```



####extractWords

```{r}
extractWords <- function(emailbody){
  uniquewords <- c()
  for (i in emailbody){
    linei <- deparse(i)
    linei <- tolower(linei)
    linei <- gsub('[[:punct:]]',' ',linei)
    linei <- gsub('[0-9]','',linei)
    linei <- gsub('\\b[A-z]{1}\\b','',linei)
    words <- lapply(strsplit(linei, " "), unique)[[1]] 
    uniquewords <- append(uniquewords,words)
    uniquewords <- uniquewords[uniquewords != ""]
  }
  return(unique(uniquewords))
}
extractWords(extractBodyText(ham424))
#extractWords(extractBodyText(readLines(spamfiles[900])))
#extractWords(extractBodyText(hardham40))
```

####readEmailDirectory

```{r}
readEmailDirectory <- function(emailpath){
  returnlist <- list()
  emailfiles <- list.files(emailpath, full.names = TRUE)
  for (i in 1:length(emailfiles)){
    returnlist[[i]] <- readLines(emailfiles[i])
  }
  return(returnlist)
}
#chubba <- head(readEmailDirectory("messages/hard_ham"))
```

####emailsAll

```{r,message=FALSE,warning=FALSE}
#creating emailsAll
hooba <- readEmailDirectory("messages/easy_ham")
hooba <- lapply(hooba, function(x) extractWords(extractBodyText(x)))
hooba2 <- readEmailDirectory("messages/hard_ham")
hooba2[[270]] <- NULL
hooba2 <- lapply(hooba2, function(x) extractWords(extractBodyText(x)))
hoobaeasyham2 <- readEmailDirectory("messages/easy_ham_2")
hoobaeasyham2 <- lapply(hoobaeasyham2, function(x) extractWords(extractBodyText(x)))
hoobaspam <- readEmailDirectory("messages/spam")
hoobaspam <- lapply(hoobaspam, function(x) extractWords(extractBodyText(x)))
hoobaspam2 <- readEmailDirectory("messages/spam_2")
hoobaspam2 <- lapply(hoobaspam2, function(x) extractWords(extractBodyText(x)))
```

```{r,warning=FALSE}
#complete emailsAll
#emailsAll[1:5050] is easy ham words
#emailsAll [5051:6450] is easy ham 2
#emailsAll[6451:6950] is hard ham
emailsAll <- c(hooba,hoobaeasyham2,hooba2,hoobaspam,hoobaspam2)
emailsAll[1:3]
```


####isSpam

```{r}
isSpam <- c()
isSpam <- rep(FALSE,6950)
isSpam <- c(isSpam,rep(TRUE,2397))
length(isSpam)
```

##Lecture 6

```{r}
#emailsTrain
emailsTrain <- emailsAll[1:4674] #ham emails
emailsTrain <- c(emailsTrain,emailsAll[6951:8508]) #spam emails


#isSpamTrain
isSpamTrain <- rep(FALSE,4674)
isSpamTrain <- c(isSpamTrain,rep(TRUE,1558))
```

```{r}
#emailsTest
emailsTest <- emailsAll[4614:6949] #ham emails
emailsTest <- c(emailsTest,emailsAll[8509:9287]) #spam emails

#isSpamTest
isSpamTest <- rep(FALSE, 2336)
isSpamTest <- c(isSpamTest,rep(TRUE,779))
```

####bag of words

```{r}
bow<-c()
for (i in 1:length(emailsTrain)){
  bow <- append(bow,emailsTrain[[i]])
  bow <- unique(bow)
}
head(bow)
```

##Lecture 7

```{R}
hamguys <- unlist(emailsTrain[1:4674])
spamguys <- unlist(emailsTrain[4675:6232])
wordsInham <- table(hamguys)
wordsInspam <- table(spamguys)
hamvector <- numeric(length(bow))
spamvector <- numeric(length(bow))
names(hamvector) <- bow
names(spamvector) <- bow
hamvector[names(wordsInham)] <- wordsInham
spamvector[names(wordsInspam)] <- wordsInspam
PresentHam <- ((hamvector+0.1)/4674.1)
PresentSpam <- (spamvector+0.1)/1558.1
AbsentHam <- 1-PresentHam
AbsentSpam <- 1-PresentSpam
logAbsentHam <- log(AbsentHam) #done
logAbsentSpam <- log(AbsentSpam)
logPresentHam <- log(PresentHam) #done
logPresentSpam <- log(PresentSpam)
```

##Lecture 8


```{r}
#computeBF
computeBF <- function(emailwords){
  emailwords <- emailwords[emailwords%in%bow] #Got rid of words not in bow
  sumpresentspam <- 0
  sumabsentspam <- 0
  sumpresentham <- 0
  sumabsentham <- 0
  for (i in emailwords){
    presentspam <- logPresentSpam[names(logPresentSpam) == i]
    sumpresentspam <- sumpresentspam + presentspam
    absentspam <- logAbsentSpam[names(logAbsentSpam) == i]
    sumabsentspam <- sumabsentspam + absentspam
    presentham <- logPresentHam[names(logPresentHam) == i]
    sumpresentham <- sumpresentham + presentham
    absentham <- logAbsentHam[names(logAbsentHam) == i]
    sumabsentham <- sumabsentham + absentham
  }
  bf <- (sumpresentspam + sumabsentspam - sumpresentham - sumabsentham)
  bf <- unname(bf)
  return(bf)
}
computeBF(emailsTest[[2]])
```

```{r}
computeBF3 <- function(emailwords){
  emailwordsnew <- emailwords[emailwords%in%bow] #Got rid of words not in bow
  sumpresentspam <- sum(logPresentSpam[names(logPresentSpam) == emailwordsnew])
  sumpresentham <- sum(logPresentHam[names(logPresentHam) == emailwordsnew])
}
```


```{R,warning=FALSE}
#computeBF2 : New approach
computeBF2 <- function(emailwords){
  emailwordsnew <- emailwords[emailwords%in%bow] #Got rid of words not in bow
  sumpresentspam <- sum(logPresentSpam[names(logPresentSpam) == emailwordsnew])
  sumpresentham <- sum(logPresentHam[names(logPresentHam) == emailwordsnew])
  bowWordsnotinEmail <- bow[bow != emailwordsnew]
  sumabsentspam <- sum(logAbsentSpam[names(logAbsentSpam) == bowWordsnotinEmail])
  sumabsentham <- sum(logAbsentHam[names(logAbsentHam) == bowWordsnotinEmail])
  bf <- (sumpresentspam+sumabsentspam-sumpresentham-sumabsentham)
  return(bf)
}
computeBF2(emailsTest[[2399]])
```

##Lecture 9

```{r,warning=FALSE}
system.time(typeIBF <- sapply(emailsTest[1:2336], function(x) computeBF2(x)))
```

```{r,warning=FALSE}
system.time(typeIIBF <- sapply(emailsTest[2337:3115], function(x) computeBF2(x)))
```

```{r}
xers <- unique(typeIBF)
yapple <- c()
for (i in xers){
  yapple <- append(yapple, (sum(typeIBF>i))/(2336*25))
}
xers2 <- unique(typeIIBF)
yapple2 <- c()
for (i in xers2){
  yapple2 <- append(yapple2, (sum(typeIIBF<i))/(779*25))
}

```

##Deliverable 9

```{r}
BFmatches <- c()
for(i in 1:length(xers)){
  if (xers[i] %in% xers2){
    BFmatches <- append(BFmatches,xers[i])
  }
}
for (i in BFmatches){
  lowerboundyapple2 <- yapple2[which(xers2==i)]-0.0015
  upperboundyapple2 <- yapple2[which(xers2==i)]+0.0015
  if(yapple[which(xers==i)] >= lowerboundyapple2 & yapple[which(xers==i)] <= upperboundyapple2){
    deliv9a <- i
    deliv9b <- (yapple[which(xers==i)])
  }
}
deliv9a
deliv9b
```

##Deliverable 10

```{r}
type1errorsmall <- xers[which(yapple<0.001)]
for (i in type1errorsmall){
  if (i %in% xers2){
    print(yapple2[which(xers2 == i)])
  }
}
```


#Deliverables

##Deliverable 1

```{r}
sapply(directory, function(x) length(list.files(paste("messages/",x,sep=''))))
```

##Deliverable 2

####a.

```{r,warning=FALSE}
hardhamfiles <- list.files("messages/hard_ham",full.names=TRUE)
deliv2attachcounter <- 0
for (i in 1:length(hardhamfiles)){
  file <- readLines(hardhamfiles[i])
  header <- splitMessage(file)$header
  if (hasAttachment(header=header)){
    deliv2attachcounter <- deliv2attachcounter+1
  }
  
}
print(deliv2attachcounter)
```

####b.

```{r}
hardham70 <- readLines(hardhamfiles[70])
getBoundary(splitMessage(hardham70)$header)
```

##Deliverable 3
```{r,warning=FALSE}
easyham2files <- list.files("messages/easy_ham_2", full.names = TRUE)
spam2files <- list.files("messages/spam_2", full.names = TRUE)
easyhamlengths <- sapply(easyhamfiles, function(x) {sum(lengths(gregexpr("\\W+",extractBodyText(readLines(x)))))+1})
spamlengths <-  sapply(spamfiles, function(x) {sum(lengths(gregexpr("\\W+",extractBodyText(readLines(x)))))+1})
easyham2lengths <- sapply(easyham2files, function(x) {sum(lengths(gregexpr("\\W+",extractBodyText(readLines(x)))))+1})
spam2lengths <- sapply(spam2files, function(x) {sum(lengths(gregexpr("\\W+",extractBodyText(readLines(x)))))+1})
#body <- extractBodyText(readLines(easyhamfiles[25]))
```

```{r,warning=FALSE}
hardhamlengths <- c()
for (i in 1:269){
  value <- sum(lengths(gregexpr("\\W+",extractBodyText(readLines(hardhamfiles[i])))))+1
  hardhamlengths <- append(hardhamlengths,value)
}
for (i in 271:length(hardhamfiles)){
  value <- sum(lengths(gregexpr("\\W+",extractBodyText(readLines(hardhamfiles[i])))))+1
  hardhamlengths <- append(hardhamlengths,value)
}
```



```{r}
boxplot(log(easyhamlengths),log(easyham2lengths),log(hardhamlengths),log(spamlengths),log(spam2lengths), names= c("easy_ham","easy_ham_2","hard_ham","spam","spam_2"),ylab="log number of words", main="Log number of words in body by email directory")
```

It seems that in general, the hard ham directory has the most number of words.

##Deliverable 4

```{r}
hamwordcount <- c()
for (i in 1:6950){
  count <- length(emailsAll[[i]])
  hamwordcount <- append(hamwordcount,count)
}
quantile(hamwordcount,0.99)
hamwordcount <- hamwordcount[which(hamwordcount<1000)]
```

```{r}
spamwordcount <- c()
for (i in 6951:9347){
  count <- length(emailsAll[[i]])
  spamwordcount <- append(spamwordcount,count)
}
spamwordcount <- spamwordcount[which(spamwordcount<1000)]

```

```{r}
hist(hamwordcount,freq = FALSE,xlim= c(0,1000),col=rgb(1,0,0,0.5),breaks=20,main="Word Count in Ham and Spam Emails", xlab= "Words")
hist(spamwordcount,freq=FALSE,col=rgb(0,0,1,0.5),xlim= c(0,1000),add=TRUE,breaks=20)
legend("topright",legend= c("ham","spam","both"),fill= c(rgb(1,0,0,0.5),rgb(0,0,1,0.5),"purple"))
```

##Deliverable 6

a.

```{r}
unname((PresentHam[names(PresentHam) == "monday"])/(PresentSpam[names(PresentSpam) == "monday"]))
```

b.

```{r}
#(PresentSpam[names(PresentSpam) == "buy"])
unname((PresentSpam[names(PresentSpam) == "buy"])/(PresentHam[names(PresentHam) == "buy"]))
```

##Deliverable 7

```{r}
boxplot(typeIBF ,typeIIBF,names=c("Ham Emails","Spam Emails"),main="Bayes Factors for Ham and Spam Emails", ylab = "Bayes Factor")
```

##Deliverable 9

```{r}
yapple[yapple == yapple2]
```

##Extensions

####Improving extractWords

```{r}
extractWords2 <- function(emailbody){
  uniquewords <- c()
  commonwords <- c("the","of","to","and","in","is","it","you","that","he","was","for","on","are","with","as","they","from","or","had","by","but","can","all","your","when","use","an","do","if","so","her","has","my","no","get","us","off","them","their","she","which","how","up","very","too","set","some","why","be")
  emailbody <- gsub("\\S*@\\S*",'',emailbody)
  for (i in emailbody){
    linei <- deparse(i)
    linei <- tolower(linei)
    linei <- gsub("http[s]?://(?:[a-zA-Z]|[0-9]|[$-_@.&+]|[!*\\(\\),]|(?:%[0-9a-fA-F][0-9a-fA-F]))+",'',linei)
    linei <- gsub('[[:punct:]]',' ',linei)
    linei <- gsub('[0-9]','',linei)
    linei <- gsub('\\b[A-z]{1}\\b','',linei)
    words <- lapply(strsplit(linei, " "), unique)[[1]] 
    uniquewords <- append(uniquewords,words)
    uniquewords <- uniquewords[uniquewords != ""]
  }
  uniquewords <- uniquewords[!(uniquewords %in% commonwords)]
  return(unique(uniquewords))
}
```


##Logistic Regression Extension

####Using the English Dictionary

```{r}
EnglishDict <- read.delim("words_alpha.txt")
EnglishDict <- EnglishDict$a
EnglishDict <- as.character(EnglishDict)
```

```{r}
length(emailsAll[[6000]][!(emailsAll[[6000]] %in% EnglishDict)])
```

```{r}
notinEnglish <- function(wordvec){
  nonEnglishwords <- length(wordvec[!(wordvec %in% EnglishDict)])
  return(nonEnglishwords)
}
```

```{r}
#notEnglishwords
notEnglish <- sapply(emailsAll, function(x) notinEnglish(x))
length(notEnglish)
```

```{r}
notEnglish[9010]
```

####What type of email is the email from?

```{r}
urlendings <- c(".com",".org",".edu",".net")
hasbadending <- function(email){
  email <- readLines(email)
  yesno <- length(grep(paste(urlendings,collapse="|"),email[1]))
  if (yesno == 1){
    valuereturn <- 0
  }
  if (yesno == 0){
    valuereturn <- 1
  }
  return(valuereturn)
}
```

```{r}
#Creating the column of weird email
badendingeasyham <- sapply(easyhamfiles, function(x) hasbadending(x))
badendingeasyham2 <- sapply(easy2hamfiles, function(x) hasbadending(x))
badendinghardham <- sapply(hardhamfiles[1:269], function(x) hasbadending(x))
hardhamassa <- sapply(hardhamfiles[271:length(hardhamfiles)], function(x) hasbadending(x))
badendinghardham <- c(badendinghardham,hardhamassa)
badendingspam <- sapply(spamfiles, function(x) hasbadending(x))
badendingspam2 <- sapply(spam2files, function(x) hasbadending(x))
badendings <- c(badendingeasyham,badendingeasyham2,badendinghardham,badendingspam,badendingspam2)
badendings <- unname(badendings)
length(badendings)
```

```{r}
#word length column vector
wordlengthcol <- c(easyhamlengths,easyham2lengths,hardhamlengths,spamlengths,spam2lengths)
```


```{r,warning=FALSE,message=FALSE}
#hasAttachment column vector
hasAttachment(splitMessage(readLines(spamfiles[900]))$header)
allemailstogether <- c(easyhamfiles,easy2hamfiles,hardhamfiles,spamfiles,spam2files)
Atteasyham <- sapply(easyhamfiles, function(x) hasAttachment(splitMessage(readLines(x))$header))
Atteasy2ham <- sapply(easy2hamfiles, function(x) hasAttachment(splitMessage(readLines(x))$header))
hasAtthardham <- sapply(hardhamfiles[1:269], function(x) hasAttachment(splitMessage(readLines(x))$header))
has22 <- sapply(hardhamfiles[271:length(hardhamfiles)], function(x) hasAttachment(splitMessage(readLines(x))$header))
hasAtthardham <- c(hasAtthardham,has22)
hasAttSpams <- sapply(c(spamfiles,spam2files), function(x) hasAttachment(splitMessage(readLines(x))$header))
#hasAttvector <- sapply(allemailstogether,function(x) hasAttachment(splitMessage(readLines(x))$header))
hasAttvector <- c(Atteasyham,Atteasy2ham,hasAtthardham,hasAttSpams)
```

```{r}
#count number of uppercase letters
library(stringr)
```

```{r}
plotErrorRates <- function(cRange){
  totalHam <- length(typeIBF)
  totalSpam <- length(typeIIBF)
  missClassifyHam <- sapply(cRange,function(x) length(typeIBF[which(typeIBF > x)]))
  missClassifySpam <- sapply(cRange, function(x) length(typeIIBF[which(typeIIBF < x)]))
  typeIerrorRate <- sapply(missClassifyHam, function(x) (x/(totalHam)))
  typeIIerrorRate <- sapply(missClassifySpam, function(x) (x/(totalSpam)))
  typeI <- data.frame(typeIerrorRate,cRange)
  typeII <- data.frame(typeIIerrorRate,cRange)
  
  maxError <- max(c(typeIerrorRate,typeIIerrorRate))
  minError <- min(c(typeIerrorRate,typeIIerrorRate))
  #typeI[which(typeI[,2]>0),]$typeIerrorRate <- rep(0,100)
  #for (i in which(typeI[,2]>0)){
  #  typeI[,1][i] <- 0.46
  #}
  for (i in 1:nrow(typeI)){
    if (typeI[i,]$cRange >94){
      typeI[i,]$typeIerrorRate <- 0.46 
    }
  }
  plot(typeI[,2]-94,typeI[,1]-0.46,xlim=c(min(cRange),max(cRange)), ylim = c(minError,maxError),col='red',type='l')
  lines(typeII[,2]+10,typeII[,1],col='green')
  lines(6:100,rep(0,95),col='red')
  return(typeI)
  #return(typeI[,1][102])
}
plotErrorRates(-75:100)
```
